\section{Compiling}

\begin{enumerate}

\item {\em Compiling fails giving me a cryptic message about a module not
  being found, usually {\tt bl\_types} or {\tt bl\_error\_module}, like:
\begin{verbatim}
mpif90 -fno-range-check -fno-second-underscore
 -Jo/3d.Linux.gcc.gfortran.MPI.EXE -I o/3d.Linux.gcc.gfortran.MPI.EXE
 -ffixed-line-length-0 -g -O3
 -I. -I/home/zingale/development/Microphysics/util
 -I../../Microphysics/EOS -I../../Microphysics/EOS/gamma_law
 -I../../Microphysics/networks
 -I../../Microphysics/networks/general_null
 -I. -I/home/zingale/development/BoxLib//Src/C_BaseLib
 -I/home/zingale/development/BoxLib//Src/C_AMRLib
 -I/home/zingale/development/BoxLib//Src/C_BoundaryLib -I../../Source
 -I../../Source/Src_3d -I../../Source/Src_nd -I../../constants
 -I/home/zingale/development/BoxLib//Src/F_BaseLib
 -I/home/zingale/development/BoxLib//Tools/C_scripts -c
 ../../Microphysics/EOS/eos.f90 -o
 o/3d.Linux.gcc.gfortran.MPI.EXE/eos.o
../../Microphysics/EOS/eos.f90:3:6:

   use bl_types
      1
Fatal Error: Can't open module file ‘bl_types.mod’ for reading at (1):
 No such file or directory compilation terminated.

/home/zingale/development/BoxLib//Tools/C_mk/Make.rules:122: recipe
 for target 'o/3d.Linux.gcc.gfortran.MPI.EXE/eos.o' failed
\end{verbatim}
}

This usually indicates that the build system cannot find a source file
(note: the problem above is not {\tt bl\_types}, that just seems to be
the way the error manifests itself).  The source files are specified
in the various {\tt Make.package} files throughout the
\castro\ directory hierarchy.  {\tt make} will look through the
directories in the {\tt VPATH\_LOCATIONS} to find the files.

There are 2 things you can do to check what's happening.  First, inspect
the directories in {\tt VPATH\_LOCATIONS}.  This can be done via:
\begin{verbatim}
make print-VPATH_LOCATIONS
\end{verbatim}

Next, ask {\tt make} to tell you where it is finding each of the source
files.  This is done through a script {\tt find\_files\_vpath.py}
 that is hooked into \castro's build system.  You can run this as:
\begin{verbatim}
make file_locations
\end{verbatim}
At the end of the report, it will list any files it cannot find in
the {\tt vpath}.  Some of these are to be expected (like {\tt extern.f90}
and {\tt buildInfo.cpp}---these are written at compiletime.  But any
other missing files need to be investigated.

\item {\em I'm still having trouble compiling.  How can I find out what
  all of the make variables are set to?}

  Use:
\begin{verbatim}
make help
\end{verbatim}
  This will tell you the value of all the compilers and their options.


\item {\em How do I use a system's BLAS library instead of compiling and
  linking the one that comes with the \starkiller\ microphysics?}

  To use a system's \index{BLAS}BLAS library, set the Make variable
  \makevar{USE\_SYSTEM\_BLAS} to {\tt TRUE}.  This will then look at
  the Make variable \makevar{BLAS\_LIBRARY} for the library to link
  (defaults to {\tt -lopenblas}).




\end{enumerate}

\section{Debugging}

\begin{enumerate}

\item {\em \castro\ crashes with a floating point exception---how can
  I get more information?}

  The best thing to do is to recompile the code with {\tt TEST=TRUE}
  set in the {\tt GNUmakefile}.  This will have \amrex\ catch the
  signals raised in both \cpp\ and Fortran functions.  Behind the
  scenes, this defines the {\tt BL\_TESTING} preprocessor flag, which
  will initialize memory allocated in {\tt fab}s or {\tt multifab}s to
  signaling NaNs (sNaN), and use the {\tt BLBackTrace::handler()}
  function to handle various signals raised in both \cpp\ and Fortran
  functions.  This is a Linux/UNIX capability.  This gives us a chance
  to print out backtrace information.  The signals include seg fault,
  floating point exceptions (NaNs, divided by zero and overflow), and
  interruption by the user and system.  What signals are handed to
  \amrex\ are controlled by \amrex (e.g., using interruption by the
  user, this was once used to find an MPI deadlock.)  It also includes
  the {\tt BL\_ASSERTION} statements if {\tt USE\_ASSERTION=TRUE} or
  {\tt DEBUG=TRUE}. 

  The \amrex\ parameters that affect the behavior are:
  \begin{itemize}
    \item {\tt amrex.fpe\_trap\_invalid}
    \item {\tt amrex.fpe\_trap\_zero}
    \item {\tt amrex.fpe\_trap\_overflow}
  \end{itemize}

  For further capabilities, defining {\tt BACKTRACE=TRUE} enables you
  to get more information than the backtrace of the call stack info by
  instrumenting the code.  (This is in \cpp\ code only). Here is an
  example.  You know the line ``{\tt Real rho = state(cell,0);}'' is
  causing a segfault.  You could add a print statement before that.
  But it might print out thousands (or even millions) of line before
  it hits the segfault.  With {\tt BACKTRACE}, you could do


\begin{verbatim}
      #ifdef BL_BACKTRACING
         std::ostringstream ss;
         ss << ``state.box() = `` << state.box() << `` cell = `` << cell;
         BL_BACKTRACE_PUSH(ss.str()); // PUSH takes std::string
      #endif

      Real rho = state(cell,0);  // state is a Fab, and cell is an IntVect.
\end{verbatim}          

    The ``print'' prints to a stack of string, not stdout.  When it
    hits the segfault, you will only see the last print out.



\item {\em How can I monitor the state in a zone from the \cpp\ side
  at various points in the evolution?}

  Given a \multifab\ {\tt mf}, you can dump out the state as:
\begin{verbatim}
    print_state(mf, IntVect(D_DECL(10, 20, 30)));
\end{verbatim}
  Here, the {\tt IntVect} has the dimension that we were compiled with
  (and this is handled through the preprocessor {\tt D\_DECL}).  In
  this case, we are inspecting zone (10, 20, 30), in the global index
  space.  Note that since a multifab exists only on a single level, the
  integer indices here refer to the global index space on that level.


\item {\em What if I want to see all the data in a \farraybox?}

  You can simply output a FAB to {\tt std::cout}.  Imagine that you
  are in an \mfiter\ loop, with a \multifab\ {\tt mf}:
\begin{verbatim}
    S = FArrayBox& mf[mfi];
    std::cout << S << std::endl;
\end{verbatim}
  This will output the contents on the FAB, one zone per line.

\end{enumerate}


\section{Profiling}

\begin{enumerate}

\item {\em How can I get line-by-line profiling information?}

  With the GNU compliers, you can enabling profiling with {\tt gprof}
  by compiling with
\begin{verbatim}
  USE_GPROF=TRUE
\end{verbatim}
  in your {\tt GNUmakefile}.

  When you run, a file named {\tt gmon.out} will be produced.  This can
  be processed with {\tt gprof} by running:
\begin{verbatim}
  gprof exec-name
\end{verbatim}
  where {\em exec-name} is the name of the executable.  More detailed
  line-by-line information can be obtained by passing the {\tt -l}
  argument to {\tt gprof}.

\end{enumerate}


\section{Managing Runs}

\begin{enumerate}

\item {\em How can I force the running code to output, even it the plot or
 checkpoint interval parameters don't require it?}

Create a file called \code{dump\_and\_continue}, e.g., as:
\begin{verbatim}
touch dump_and_continue
\end{verbatim}

This will force the code to output a checkpoint file that can be used
to restart.  Other options are \code{plot\_and\_continue} to output
a plotfile, \code{dump\_and\_stop} to output a checkpoint file
and halt the code, and \code{stop\_run} to simply stop the code.
Note that the parameter \runparam{amr.message\_int} controls how often
the existence of these files is checked; by default it is 10, so the
check will be done at the end of every timestep that is a multiple of 10.
Set that to 1 in your inputs file if you'd like it to check every timestep.

\end{enumerate}

\section{Visualization}

\label{ch:faq:vis}

\begin{enumerate}

\item {\em When I try to use \amrvis\ with the Nvidia driver, all I see is
  black---no data.  How do I fix this?}

  You need to edit your {\tt xorg.conf} file (usually found in {\tt /etc/X11/}
to enable the {\tt Dac8Bit} option.  The section will look like:
\begin{verbatim}
Section "Device"
    Identifier     "Device0"
    Driver         "nvidia"
    VendorName     "NVIDIA Corporation"
    Option         "Dac8bit" "True"
EndSection
\end{verbatim}
If you don't already have an {\tt xorg.conf} then you can create one
by running {\tt nvidia-xconfig} first.


\end{enumerate}
