\subsection{List of Parameters}

Note that in the following, {\tt MAX\_LEV} is a hard-coded
parameter in Source/Gravity.cpp which is currently set to 15.
It determines how many levels can be tracked by the Gravity object.

\begin{table*}[h]
\begin{scriptsize}
\begin{center}
\begin{tabular}{|l|l|l|l|} \hline
Parameter & Definition & Acceptable Values &Default\\
\hline
{\bf castro.do\_grav}  & Include gravity as a forcing term & 0 if false, 1 if true & must be set \\
& & & if USE\_GRAV = TRUE  \\
\hline
{\bf gravity.gravity\_type} & if {\bf castro.do\_grav} = 1, & ConstantGrav, & \\
& how shall gravity be calculated & PoissonGrav, & \\
& & or MonopoleGrav & must be set \\
\hline
{\bf gravity.const\_grav} & if {\bf gravity.gravity\_type} = ConstantGrav, & & \\
& set the value of constant gravity & Real & 0.0 \\
\hline
{\bf gravity.no\_sync} & if {\bf gravity.gravity\_type} = PoissonGrav, & & \\
& whether to perform the ``sync solve" &  0 or 1 & 0 \\
\hline
{\bf gravity.no\_composite} & if {\bf gravity.gravity\_type} = PoissonGrav, & & \\
& whether to perform a composite solve & 0 or 1 & 0 \\
\hline
{\bf gravity.max\_solve\_level} & maximum level to solve for $\phi$ and $\mathbf{g}$; & & \\
& above this level, interpolate from below & Integer & ${\tt MAX\_LEV} - 1$ \\
\hline
{\bf gravity.ml\_tol} & if {\bf gravity.gravity\_type} = PoissonGrav, & & \\
& this is the tolerance for level solves. & & \\
& It can be either a single value for all levels & & \\
& or you can specify one for each level. & Real & 1.e-11 \\
\hline
{\bf gravity.delta\_tol} & if {\bf gravity.gravity\_type} = PoissonGrav, & & \\
& this is the tolerance for sync solves. & & \\
& It can be either a single value for all levels & & \\
& or you can specify one for each level. & Real & 1.e-11 \\
\hline
{\bf gravity.max\_multipole\_order} & if {\bf gravity.gravity\_type} = PoissonGrav, & & \\
& max $\ell$ value to use for multipole BCs & Integer $\geq 0$ & 0 \\
\hline
{\bf gravity.direct\_sum\_bcs} & if {\bf gravity.gravity\_type} = PoissonGrav, & & \\
& evaluate BCs using exact sum & 0 or 1 & 0 \\
\hline
{\bf castro.point\_mass}   & Point mass at the center of the star & Real $\geq 0$ & 0.0 \\

\hline
\end{tabular}
\label{Table:Gravity}
\end{center}
\end{scriptsize}
\end{table*}

\subsection{Notes}

    You must have USE\_POINTMASS = TRUE in the GNUmakefile for {\bf
      castro.point\_mass} to be relevant.
    
Gravity types are ConstantGrav, PoissonGrav, or MonopoleGrav.  See the Gravity section for more detail.

\begin{itemize}
\item To include gravity you must set 
\begin{itemize}
\item USE\_GRAV  = TRUE in the GNUmakefile 
\item {\bf castro.do\_grav} = 1 in the inputs file
\end{itemize}
\item {\bf gravity.gravity\_type} is  only relevant if {\bf castro.do\_grav} = 1 
\item {\bf gravity.no\_sync} and {\bf gravity.no\_composite} are only relevant if {\bf gravity.gravity\_type} = PoissonGrav,
i.e. the code does a full Poisson solve for self-gravity.
\item {\bf gravity.max\_multipole\_order} and {\bf gravity.direct\_sum\_bcs} are only relevant if {\bf gravity.gravity\_type} = PoissonGrav,
and they are described in Section \ref{sec-poisson-3d-bcs}.
\end{itemize}


There are currently four options for how gravity is calculated -- 
these are controlled by setting {\bf gravity.gravity\_type}.  
The options are {\bf ConstantGrav, PoissonGrav, Monopole Grav} or {\bf PrescribedGrav}.
Note that these are only relevant if {\bf USE\_GRAV = TRUE} in the GNUmakefile
and {\bf castro.do\_grav} = 1 in the inputs file.  If both of these are
set then the user is required to specify the gravity type in the inputs file
or the program will abort. 

Note that {\bf MonopoleGrav} and {\bf PoissonGrav} is only correct for 
spherical stars, i.e. in 1D we must have {\bf coord\_sys} = 2, 
in 2D we must have {\bf coord\_sys} = 1, and in 3D we only support {\bf coord\_sys} = 0.

\section{Types of Approximations}

\subsection{ConstantGrav}

Gravity can be defined as constant in direction and magnitude, 
defined in the inputs file by 

\noindent {\bf gravity.const\_grav} = -9.8

for example, to set the gravity to have magnitude 9.8 in the 
negative y-direction if in 2D, negative z-direction if in 3-D.

\subsection{PoissonGrav}

The most general case is a self-induced gravitational field, 
\begin{equation}
\mathbf{g}(\mathbf{x},t) = \nabla \phi 
\end{equation}
where $\phi$ is defined by solving
\begin{equation}
\mathbf{\Delta} \phi = -4 \pi G \rho .\label{eq:Self Gravity}
\end{equation}
(We note that the sign convention used for $\phi$ here is opposite
that traditionally used in astrophysics, but the resulting
gravitational acceleration will be the same.)

We only allow {\bf PoissonGrav} in 2D or 3D because in 1D, computing the monopole 
approximation in spherical coordinates is faster and more accurate than solving
the Poisson equation.

\subsubsection{Poisson Boundary Conditions: 2D}

In 2D, if boundary conditions are not periodic in both directions, we use a 
monopole approximation at the coarsest level. This involves computing an effective
1D radial density profile (on \texttt{level == 0} only), integrating it outwards
from the center to get the gravitational acceleration $\mathbf{g}$, and then integrating
$g$ outwards from the center to get $\phi$ (using $\phi(0) = 0$ as a boundary condition,
since no mass is enclosed at $r = 0$). For more details, see Section \ref{sec-monopole-grav}.

\subsubsection{Poisson Boundary Conditions: 3D}\label{sec-poisson-3d-bcs}

\noindent{\bf Note: The text here, and the multipole and direct sum boundary conditions, were 
contributed by Max Katz of Stony Brook University.}

\begin{itemize}
\item \textbf{Multipole Expansion}

In 3D, by default, we use a multipole expansion to estimate the value of the boundary conditions. According to, for example, Jackson's \textit{Classical Electrodynamics} (with the corresponding change to Poisson's equation for electric charges and gravitational ''charges''), an expansion in spherical harmonics for $\phi$ is
\begin{equation}
  \phi(\mathbf{x}) = G\sum_{l=0}^{\infty}\sum_{m=-l}^{l} \frac{4\pi}{2l + 1} q_{lm} \frac{Y_{lm}(\theta,\phi)}{r^{l+1}}, \label{spherical_harmonic_expansion}
\end{equation}
The origin of the coordinate system is taken to be the \texttt{center} variable, that must be declared and stored in the \texttt{probdata} module in your project directory. The validity of the expansion used here is based on the assumption that a sphere centered on \texttt{center}, of radius approximately equal to the size of half the domain, would enclose all of the mass. Furthermore, the lowest order terms in the expansion capture further and further departures from spherical symmetry. Therefore, it is crucial that \texttt{center} be near the center of mass of the system, for this approach to achieve good results.

The multipole moments $q_{lm}$ can be calculated by expanding the Green's function for the Poisson equation as a series of spherical harmonics, which yields
\begin{equation}
  q_{lm} = \int Y^*_{lm}(\theta^\prime, \phi^\prime)\, {r^\prime}^l \rho(\mathbf{x}^\prime)\, d^3x^\prime. \label{multipole_moments_original}
\end{equation}
Some simplification of Equation \ref{spherical_harmonic_expansion} can be achieved by using the addition theorem for spherical harmonics:
\begin{align}
  &\frac{4\pi}{2l+1} \sum_{m=-l}^{l} Y^*_{lm}(\theta^\prime,\phi^\prime)\, Y_{lm}(\theta, \phi) = P_l(\text{cos}\, \theta) P_l(\text{cos}\, \theta^\prime) \notag \\
  &\ \ + 2 \sum_{m=1}^{l} \frac{(l-m)!}{(l+m)!} P_{l}^{m}(\text{cos}\, \theta)\, P_{l}^{m}(\text{cos}\, \theta^\prime)\, \left[\text{cos}(m\phi)\, \text{cos}(m\phi^\prime) + \text{sin}(m\phi)\, \text{sin}(m\phi^\prime)\right].
\end{align}
Here the $P_{l}^{m}$ are the associated Legendre polynomials and the $P_l$ are the Legendre polynomials. After some algebraic simplification, the potential outside of the mass distribution can be written in the following way:
\begin{equation}
  \Phi(\mathbf{x}) \approx G\sum_{l=0}^{l_{\text{max}}} \left[Q_l^{(0)} \frac{P_l(\text{cos}\, \theta)}{r^{l+1}} + \sum_{m = 1}^{l}\left[ Q_{lm}^{(C)}\, \text{cos}(m\phi) + Q_{lm}^{(S)}\, \text{sin}(m\phi)\right] \frac{P_{l}^{m}(\text{cos}\, \theta)}{r^{l+1}} \right].
\end{equation}
The modified multipole moments are:
\begin{align}
  Q_l^{(0)}   &= \int P_l(\text{cos}\, \theta^\prime)\, {r^{\prime}}^l \rho(\mathbf{x}^\prime)\, d^3 x^\prime \\
  Q_{lm}^{(C)} &= 2\frac{(l-m)!}{(l+m)!} \int P_{l}^{m}(\text{cos}\, \theta^\prime)\, \text{cos}(m\phi^\prime)\, {r^\prime}^l \rho(\mathbf{x}^\prime)\, d^3 x^\prime \\
  Q_{lm}^{(S)} &= 2\frac{(l-m)!}{(l+m)!} \int P_{l}^{m}(\text{cos}\, \theta^\prime)\, \text{sin}(m\phi^\prime)\, {r^\prime}^l \rho(\mathbf{x}^\prime)\, d^3 x^\prime.
\end{align}
Our strategy for the multipole boundary conditions, then, is to pick some value $l_{\text{max}}$ that is of sufficiently high order to capture the distribution of mass on the grid, evaluate the discretized analog of the modified multipole moments for $0 \leq l \leq l_{\text{max}}$ and $1 \leq m \leq l$, and then directly compute the value of the potential on all of the boundary zones. This is ultimately an $\mathcal{O}(N^3)$ operation, the same order as the monopole approximation, and the wall time required to calculate the boundary conditions will depend on the chosen value of $l_{\text{max}}$.

The number of $l$ values calculated is controlled by \texttt{gravity.max\_multipole\_order} in your inputs file. By default, it is set to \texttt{0}, which means that a monopole approximation is used. There is currently a hard-coded limit of $l_{\text{max}} = 50$. This is because the method used to generate the Legendre polynomials is not numerically stable for arbitrary $l$ (because the polynomials get very large, for large enough $l$).

\item \textbf{Direct Sum}

Up to truncation error caused by the discretization itself, the boundary values for the potential can be computed exactly by a direct sum over all cells in the grid. Suppose I consider some ghost cell outside of the grid, at location $\mathbf{r}^\prime \equiv (x^\prime, y^\prime, z^\prime)$. By the principle of linear superposition as applied to the gravitational potential,
\begin{equation}
  \phi(\mathbf{r}^\prime) = \sum_{\text{ijk}} \frac{G \rho_{\text{ijk}}\, \Delta V_{\text{ijk}}}{\left[(x - x^\prime)^2 + (y - y^\prime)^2 + (z - z^\prime)^2\right]^{1/2}},
\end{equation}
where $x = x(i)$, $y = y(j)$ and $z = z(k)$ are constructed in the usual sense from the zone indices. The sum here runs over every cell in the physical domain (that is, the calculation is $\mathcal{O}(N^3)$ for each boundary cell). There are $6N^2$ ghost cells needed for the Poisson solve (since there are six physical faces of the domain), so the total cost of this operation is $\mathcal{O}(N^5)$ (this only operates on the coarse grid, at present). In practice, we use the domain decomposition inherent in the code to implement this solve: for the grids living on any MPI task, we create six $N^2$ arrays representing each of those faces, and then iterate over every cell on each of those grids, and compute their respective contributions to all of the faces. Then, we do a global reduce to add up the contributions from all cells together. Finally, we place the boundary condition terms appropriate for each grid onto its respective cells.

This is quite expensive even for reasonable sized domains, so this option is recommended only for analysis purposes, to check if the other methods are producing accurate results. It can be enabled by setting \texttt{gravity.direct\_sum\_bcs = 1} in your inputs file.

\end{itemize}

\subsection{MonopoleGrav}\label{sec-monopole-grav}
\begin{itemize}

\item In 1D spherical coordinates we compute 
\[g(r) = G * ({\rm Mass_{enclosed}}) / r^2 \; \; , \]
where ${\rm Mass_{enclosed}}$ is calculated from the density at the time of the call. 
For levels above the coarsest level we define the extent of that level's radial arrays 
as ranging from the center of the star ($r=0$) to the cell at that level farthest away from the
origin.  If there are gaps between fine grids in that range then we interpolate the density
from a coarser level in order to construct a continuous density profile.  We note that
the location of values in the density profile and in the gravitational field exactly match the
location of data at that level so there is no need to interpolate between points when
mapping the 1D radial profile of $g$ back onto the original grid.

\item In 2D or 3D we compute a 1D radial average of density and use this to compute 
gravity as a one-dimensional integral, then interpolate the gravity vector back onto 
the Cartesian grid cells. At the coarsest level we define the extent of the 1D arrays as
ranging from the center of the star to the farthest possible point in the grid (plus a
few extra cells so that we can fill ghost cell values of gravity).  At finer levels
we first define a single box that contains all boxes on that fine level, then we
interpolate density from coarser levels as needed to fill the value of density at every
fine cell in that box.   The extent of the radial array is from the center of the star
to the {\em nearest} cell on one of the faces of the single box.  This ensures that
all cells at that maximum radius of the array are contained in this box.

We then average the density onto a 1D radial array.
We note that there is a mapping from the Cartesian cells to the radial array and back;
unlike the 1D case this requires interpolation. We use quadratic interpolation with
limiting so that the interpolation does not create new maxima or minima.

The default resolution of the radial arrays at a level is the grid cell spacing at
that level, i.e. $\Delta r = \Delta x.$   One optimization we have recently added is that one 
can define {\bf castro.drdxfac} as a number greater than 1 (2 or 4 are recommended) and the
spacing of the radial array will then satisfy $\Delta x / \Delta r = $ {\bf drdxfac}.
Individual Cartesian grid cells are subdivided by {\bf drdxfac} in each coordinate direction
for the purposing of averaging the density, and the integration that creates $g$ is
done at the finer resolution of the new $\Delta r.$   

Note that the center of the star is defined in the subroutine PROBINIT and the radius
is computed as the distance from that center.

\subsection{PrescribedGrav}

\noindent{\bf Note: The {\bf PrescribedGrav} option and text here were 
contributed by Jan Frederik Engels of University of Gottingen.}

With this option, gravity can be defined as a function that
is specified by the user.  The option is allowed in 2D and 3D.  To define the gravity
vector, copy {\bf prescribe\_grav\_nd.f90} from {\bf Src\_nd} to your run directory.
The makefile system will always choose this local copy of the file over the one in another directory.
Then define the components of gravity inside a loop over the grid inside the file.
If your problem uses a radial gravity in the form $g(r)$, you can simply adapt 
{\bf ca\_prescribe\_grav\_gravityprofile}, otherwise you will have to adapt 
{\bf ca\_prescribe\_grav}, both are located in {\bf prescribed\_grav\_nd.90}.
\end{itemize}

\section{GR correction}

\noindent{\bf Note: The GR code and text here were contributed by Ken Chen of Univ. of Minnesota.}

In the cases of compact objects or very massive stars, the general relativity 
(GR) effect starts to play a role. First, we consider the hydrostatic equilibrium 
due to effects of GR then derive GR-correction term for Newtonian gravity. 
The correction term is applied to the monopole approximation only when 
{\bf USE\_GR  = TRUE} is set in the GNUmakefile. 

The formulae of GR-correction here are based on \cite{grbk1}. For detailed physics,
 please refer to \cite{grbk2}. For describing very strong gravitational field, we need to use Einstein 
field equations
\begin{equation}\label{field}
R_{ik}-\frac{1}{2}g_{ik}R=\frac{\kappa}{c^{2}}T_{ik} \quad , \quad
\kappa=\frac{8\pi G}{c^{2}}\quad ,
\end{equation} 
where $R_{ik}$ is the Ricci tensor, $g_{ik}$ is the metric tensor, $R$ is the Riemann curvature, $c$ is 
the speed of light and $G$ is gravitational constant. $T_{ik}$ is the energy momentum tensor, which for 
ideal gas has only the non-vanishing components $T_{00}$ = $\varrho c^2$ , $T_{11}$ = $T_{22}$ = $T_{33}$ = $P$ 
( contains rest mass and energy density, $P$ is pressure). We are interested in spherically symmetric mass 
distribution. Then the line element $ds$ for given spherical coordinate $(r, \vartheta, \varphi)$ has the 
general form
\begin{equation}\label{metric}
  ds^{2} = e^{\nu}c^{2}dt^{2}-e^{\lambda}dr^{2}-r^{2}(d\vartheta^{2}+\sin^{2}
  \vartheta d\varphi) \quad ,
\end{equation} 
with $\nu = \nu(r)$, $\lambda = \lambda(r)$. Now we can put the expression of $T_{ik}$ and $ds$ into (\ref{field}), then field equations can be reduced to 3 ordinary differential equations:
\begin{equation}\label{diff1}
   \frac{\kappa P}{c^{2}} =
   e^{-\lambda}(\frac{\nu^{\prime}}{r}+\frac{1}{r^{2}})-\frac{1}{r^{2}}
   \quad ,
\end{equation}
\begin{equation}\label{diff2}
  \frac{\kappa P}{c^{2}} =
  \frac{1}{2}e^{-\lambda}(\nu^{\prime\prime}+\frac{1}{2}{\nu^{\prime}}^{2}+\frac{\nu^
    {\prime}-\lambda^{\prime}}{r}
   -\frac{\nu^{\prime}\lambda^{\prime}}{2}) \quad ,
\end{equation}
\begin{equation}\label{diff3}
  \kappa \varrho =
  e^{-\lambda}(\frac{\lambda^{\prime}}{r}-\frac{1}{r^{2}})+\frac{1}{r^{2}} \quad ,
\end{equation} 
where primes means the derivatives with respect to $r$. After multiplying with $4\pi r^2$, (\ref{diff3}) can 
be integrated and yields
\begin{equation}\label{gmass1}
  \kappa m = 4\pi r (1-e^{-\lambda}) \quad ,
\end{equation}
the $m$ is called ``gravitational mass'' inside r defined as
\begin{equation}\label{gmass2}
  m = \int_{0}^{r}4\pi r^{2}  \varrho dr\quad .
\end{equation}      
For the $r = R$, $m$ becomes the mass $M$ of the star. $M$ contains not only the rest mass but
the whole energy (divided by $c^2$), that includes the internal and gravitational energy. So the
 $\varrho = \varrho_0 +U/c^2$ contains the whole energy density $U$ and rest-mass density $\varrho_0$. 
Differentiation of (\ref{diff1}) with respect to $r$ gives $P = P^{\prime}(\lambda,\lambda^{\prime},
\nu,\nu^{\prime},r)$, where $\lambda,\lambda^{\prime},\nu,\nu^{\prime}$  can be eliminated by (\ref{diff1}), 
(\ref{diff2}), (\ref{diff3}). Finally we reach \textit{Tolman-Oppenheinmer-Volkoff(TOV)} equation for 
hydrostatic equilibrium in general relativity:
\begin{equation}\label{tov}
  \frac{dP}{dr} = -\frac{Gm}{r^{2}}\varrho (1+\frac{P}{\varrho
    c^{2}})(1+\frac{4\pi r^3 P}{m c^{2}}) (1-\frac{2Gm}{r c^{2}})^{-1} \quad .
\end{equation}                  
For Newtonian case $c^2 \rightarrow  \infty $, it reverts to usual form
\begin{equation}\label{newton}
  \frac{dP}{dr} = -\frac{Gm}{r^{2}}\varrho \quad .
\end{equation}
Now we take effective monopole gravity as
\begin{equation}\label{tov2}                                                      
\tilde{g} = -\frac{Gm}{r^{2}} (1+\frac{P}{\varrho
  c^{2}})(1+\frac{4\pi r^3 P}{m c^{2}}) (1-\frac{2Gm}{r c^{2}})^{-1}  \quad .
\end{equation}                                 
For general situations, we neglect the $U/c^2$ and potential energy in m because they are usually
much smaller than $\varrho_0$. Only when $T$ reaches $10^{13} K$ ($KT \approx m_{p} c^2$, $m_p$ is proton mass)
 before it really makes a difference. So (\ref{tov2}) can be expressed as
\begin{equation}\label{tov3}                                                      
  \tilde{g} = -\frac{GM_{enc}}{r^{2}} (1+\frac{P}{\varrho
    c^{2}})(1+\frac{4\pi r^3 P}{M_{enc} c^{2}}) (1-\frac{2GM_{enc}}{r c^{2}})^{-1} \quad ,
\end{equation}                                              
where $M_{enc}$ is identical to $Mass_{enclosed}$ in the previous section.
